---
title: "18S_analysis_steps"
author: "Gemma Clucas"
date: "11/20/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load qiime on laptop and ```cd``` to correct directory.

```
conda activate qiime2-2019.4
cd /Users/gemmaclucas/Dropbox/Diets_from_poop/2019_terns_puffins_fecal_data_analysis/18S
```


## Primary vs secondary consumption

This is a big issue with the 18S data. Lots of copepod sequences come up, which are definitely due to secondary consumption.

The Birds of the World account for Common Terns only reports them feeding on fish, insects, crustaceans, and squid. The account for Roseates reports even fewer non-fish prey. However, making arbitrary decisions about what may or may not be primary prey is unnecessary. If I present all non-parasite data, then I can talk about how this captures information about the food web as a whole i.e. the importance of copepods in the system. I can also use prey size and FOO to determine what are important components of the tern diets and what isn't.

## Separate out the data from different tern species

Extract the COTE sequences from the table and rep-seqs. 
```
qiime feature-table filter-samples \
  --i-table merged_table_sklearn_onlymetazoa.qza \
  --m-metadata-file mdat.txt \
  --p-where "Species='COTE'" \
  --o-filtered-table Terns/COTE_merged_table_onlymetazoa.qza 
  
qiime feature-table filter-seqs \
  --i-data merged_rep-seqs.qza \
  --i-table Terns/COTE_merged_table_onlymetazoa.qza \
  --o-filtered-data Terns/COTE_seqs_onlymetazoa.qza
```

Create the ```.qzv``` files to go with the table and rep-seqs.
```
qiime feature-table summarize \
    --i-table Terns/COTE_merged_table_onlymetazoa.qza \
    --m-sample-metadata-file mdat.txt \
    --o-visualization Terns/COTE_merged_table_onlymetazoa
    
qiime feature-table tabulate-seqs \
    --i-data Terns/COTE_seqs_onlymetazoa.qza \
    --o-visualization Terns/COTE_seqs_onlymetazoa  
    
qiime taxa barplot \
  --i-table Terns/COTE_merged_table_onlymetazoa.qza \
  --i-taxonomy sklearn_taxonomy.qza \
  --m-metadata-file mdat.txt \
  --o-visualization Terns/COTE_merged_table_onlymetazoa_barplot    
```

Extract the ROST sequences from the table and rep-seqs. 
```
qiime feature-table filter-samples \
  --i-table merged_table_sklearn_onlymetazoa.qza \
  --m-metadata-file mdat.txt \
  --p-where "Species='ROST'" \
  --o-filtered-table Terns/ROST_merged_table_onlymetazoa.qza 
  
qiime feature-table filter-seqs \
  --i-data merged_rep-seqs.qza \
  --i-table Terns/ROST_merged_table_onlymetazoa.qza \
  --o-filtered-data Terns/ROST_seqs_onlymetazoa.qza
```

Create the ```.qzv``` files to go with the table and rep-seqs.
```
qiime feature-table summarize \
    --i-table Terns/ROST_merged_table_onlymetazoa.qza \
    --m-sample-metadata-file mdat.txt \
    --o-visualization Terns/ROST_merged_table_onlymetazoa
    
qiime feature-table tabulate-seqs \
    --i-data Terns/ROST_seqs_onlymetazoa.qza \
    --o-visualization Terns/ROST_seqs_onlymetazoa 
    
qiime taxa barplot \
  --i-table Terns/ROST_merged_table_onlymetazoa.qza \
  --i-taxonomy sklearn_taxonomy.qza \
  --m-metadata-file mdat.txt \
  --o-visualization Terns/ROST_merged_table_onlymetazoa_barplot       
```

## Exclude bird, mammal, and parasite DNA from ROST data

Only done for ROST so far, need to go through the barplots from COTE to find the right terms to exlude before running these commands.

```
qiime taxa filter-table \
  --i-table Terns/ROST_merged_table_onlymetazoa.qza \
  --i-taxonomy sklearn_taxonomy.qza \
  --p-exclude Aves,Mammalia,Chromadore,Myxosporea,Acari,Digenea,Eucestoda \
  --o-filtered-table Terns/ROST_table_noBirdsMammalsParasites

qiime taxa barplot \
  --i-table Terns/ROST_table_noBirdsMammalsParasites.qza \
  --i-taxonomy sklearn_taxonomy.qza \
  --m-metadata-file mdat.txt \
  --o-visualization Terns/ROST_noBirdsMammalsParasites_barplot
```

## Check PCR and extraction blanks for contamination (all tern species)

I checked the extraction and PCR blanks for the ROST samples and all are clear or below levels that will be removed by rarefaction (fewer than 300 sequences). There is one instance (EXT_BLANK_7) where there are a large amount of sequences in the blank, but this is due to cross-contamination from an adjacent well, as noted in my lab notebook at the time. Therefore, I do not need to account for cross-contamination in my analysis of the ROST samples. 

Field blanks were totally clear i.e. no sequences reached this stage in the analysis.

**I should remove ROST_SEA_CH_496 and ROST_SEA_CH_498 because they could also have been affected by cross-contamination**

## Calculate rarefaction depth for ROST data

I need to collapse taxonomy first, so I am not counting ASVs. This does make a difference to the rarefaction curves calculated below.

```
qiime taxa collapse \
    --i-table Terns/ROST_table_noBirdsMammalsParasites.qza \
    --i-taxonomy sklearn_taxonomy.qza \
    --p-level 7 \
    --o-collapsed-table Terns/ROST_collapsedtable_noBirdsMammalsParasites.qza
```

Calculate rarefection curves broadly up to 20,000, then zoom in on the left hand side of the plot to figure out the exact cut-off to use.
```
qiime diversity alpha-rarefaction \
  --i-table Terns/ROST_collapsedtable_noBirdsMammalsParasites.qza \
  --m-metadata-file mdat.txt \
  --p-min-depth 100 \
  --p-max-depth 20000 \
  --o-visualization Terns/ROST_noBirdsMammalsParasite_alpha-rarefaction_100-20000
  
qiime diversity alpha-rarefaction \
  --i-table Terns/ROST_collapsedtable_noBirdsMammalsParasites.qza \
  --m-metadata-file mdat.txt \
  --p-min-depth 100 \
  --p-max-depth 6000 \
  --o-visualization Terns/ROST_noBirdsMammalsParasite_alpha-rarefaction_100-6000
  
```
The number of observed OTUs is flat at 4 from 750 to 4750 reads, but then increases to five when you reach 6000 reads. Given I have barely any samples with more than 6000 sequences, I will rarefy to 750 for now but bear this in mind for writing up the data.


## Rarefy to a sampling depth of 750 for ROST

Note, this needs to be done on the original feature table, not the collapsed one.

```
qiime feature-table rarefy \
  --i-table Terns/ROST_table_noBirdsMammalsParasites.qza \
  --p-sampling-depth 750 \
  --o-rarefied-table Terns/ROST_table_rarefied750 
```

Redo the barplots for the rarefied data
```
qiime taxa barplot\
      --i-table Terns/ROST_table_rarefied750.qza \
      --i-taxonomy sklearn_taxonomy.qza \
      --m-metadata-file mdat.txt \
      --o-visualization Terns/ROST_table_rarefied750-barplots.qzv
```

Then I exported the ```.csv``` file from Qiime Viewer using level 6 taxonomy and grouped all Copepoda (the vast majority were Calanoida).

## Create FOO barplot

```{r}
library(ggplot2)
library(RColorBrewer)
library(reshape2)
library(tidyverse)
library(viridis)
library(knitr)

df <- read.csv("/Users/gemmaclucas/Dropbox/Diets_from_poop/2019_terns_puffins_fecal_data_analysis/18S/Terns/ROST_table_rarefied750_level6_namesEdited.csv", header = TRUE) %>% 
  mutate(Year = as.factor(Year))
```

I need to define the order of the taxa along the x-axis, so that it is consistent among graphs.

```{r}

Species_ordered <- c("Teleostei",
             "Copepoda",
             "Eucarida",
             "Hymenoptera",
             "Hemiptera",
             "Heterobranchia",
             "Collembola",
             "Lepidoptera",
             "Coleoptera",
             "Neoptera",
             "Araneae",
             "Caenogastropoda",
             "Semaeostomeae",
             "Mytiloida",
             "Ploimida")

Species_with_colours <- c("Teleostei"= rgb(98, 64, 18, max = 255),
             "Copepoda" = rgb(150, 102, 34, max = 255),
             "Eucarida" = rgb(207, 138, 54, max = 255),
             "Hymenoptera" = rgb(247, 224, 167, max = 255),
             "Hemiptera" = rgb(225, 204, 150, max = 255),
             "Heterobranchia" = rgb(187, 169, 126, max = 255),
             "Collembola" = rgb(214,	237, 234, max = 255),
             "Lepidoptera" = rgb(159, 211, 204, max = 255),
             "Coleoptera" = rgb(92,	164, 160, max = 255),
             "Neoptera" = rgb(50,	118, 113, max = 255),
             "Araneae" = rgb(30, 75, 63, max = 255),
             "Caenogastropoda" = rgb(5, 60, 245, max = 255),
             "Semaeostomeae" = rgb(5, 57, 237, max = 255),
             "Mytiloida" = rgb(6, 56, 234, max = 255),
             "Ploimida" = rgb(4, 50, 211, max = 255))


```

Functions to calculate FOO and plot
```{r}

calc_FOO <- function(x) {
  x %>% 
  mutate_if(is.numeric, ~1 * (. > 0)) %>%     # change to detection/non-detection
  summarise_each(funs = sum) %>%              # count number of detections
  melt() %>%                                  # make into long dataframe
  rename(Occurrence = value,
         Species = variable) %>% 
  mutate(FOO = Occurrence/n_samples*100)
}


# get rid of periods in column names and order species
scrub_periods <- function(x) {
  x$Species <-  gsub("\\.", " ", x$Species)
  x$Species <- factor(x$Species, levels = Species_ordered)
}

# function to plot, with the species in the order determined by Species_with_colours
plot_FOO <- function(x) {
  x %>% 
    ggplot() +
    geom_bar(aes(x = Species, y = FOO, fill = Species), stat = "identity") +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45,  hjust=1)) +
    labs(y = "Frequency of occurrence (%)") 
    #scale_fill_manual(values = Species_with_colours)
}


```

Select the data for each year and age class
```{r}
# Filter the data and save to object
ROST_chicks_2017 <- df %>% 
  filter(Age == "chick" & Year == 2017) %>% 
  select(Coleoptera:Eucarida) 
n_samples <- nrow(ROST_chicks_2017)
FOO_ROST_chicks_2017 <- calc_FOO(ROST_chicks_2017)
FOO_ROST_chicks_2017$Year <- "2017"
FOO_ROST_chicks_2017$Age <- "chick"

# Filter the data and save to object
ROST_chicks_2018 <- df %>% 
  filter(Age == "chick" & Year == 2018) %>% 
  select(Coleoptera:Eucarida) 
n_samples <- nrow(ROST_chicks_2018)
FOO_ROST_chicks_2018 <- calc_FOO(ROST_chicks_2018)
FOO_ROST_chicks_2018$Year <- "2018"
FOO_ROST_chicks_2018$Age <- "chick"

# Filter the data and save to object
ROST_adults_2018 <- df %>% 
  filter(Age == "adult" & Year == 2018) %>% 
  select(Coleoptera:Eucarida) 
n_samples <- nrow(ROST_adults_2018)
FOO_ROST_adults_2018 <- calc_FOO(ROST_adults_2018)
FOO_ROST_adults_2018$Year <- "2018"
FOO_ROST_adults_2018$Age <- "adult"
```

Use row binding to add all the dataframes together so that I can use faceting to plot.

```{r}
All_FOO <- FOO_ROST_chicks_2017 %>% 
  bind_rows(., FOO_ROST_chicks_2018) %>% 
  bind_rows(., FOO_ROST_adults_2018)

All_FOO$Species <- scrub_periods(All_FOO)

kable(All_FOO)

```



```{r}
p <- All_FOO %>% 
  group_by(Species) %>% 
  filter(FOO > 0) %>% 
  ggplot() +
  geom_bar(aes(x = Species, y = FOO, fill = Species), stat = "identity") +
  facet_grid(rows = vars(Year), cols = vars(Age)) +
  ylim(0, 100) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45,  hjust=1, colour = "black", margin = margin(t = 0)),
        axis.text.y = element_text(colour = "black"),
        axis.text=element_text(size=7),
        axis.title=element_text(size=8, face = "bold"),
        axis.ticks.x = element_blank(),
        panel.border=element_rect(colour="black",size=1, fill = NA),
        panel.spacing = unit(2, "mm"),
        strip.background = element_blank(),
        strip.text = element_text(colour = "black", size=8, face = "bold"),
        axis.line = element_blank()) +
  labs(y = "Frequency of occurrence (%)") +
  scale_fill_manual(values = Species_with_colours) +
  theme(legend.position = "none")

p
```

```{r}
q <- All_FOO %>% 
  group_by(Species) %>% 
  filter(FOO > 0) %>% 
  ggplot() +
  geom_bar(aes(x = Species, y = FOO, fill = Species), stat = "identity") +
  facet_grid(rows = vars(Year)) +
  ylim(0, 100) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45,  hjust=1, colour = "black", margin = margin(t = 0)),
        axis.text.y = element_text(colour = "black"),
        axis.text=element_text(size=7),
        axis.title=element_text(size=8, face = "bold"),
        axis.ticks.x = element_blank(),
        panel.border=element_rect(colour="black",size=1, fill = NA),
        panel.spacing = unit(2, "mm"),
        strip.background = element_blank(),
        strip.text = element_text(colour = "black", size=8, face = "bold"),
        axis.line = element_blank()) +
  labs(y = "Frequency of occurrence (%)") +
  #scale_fill_manual(values = Species_with_colours) +  # colour scale Liz used
  scale_fill_manual(values = c(rep.int(rgb(182, 52, 19, max = 255), 1), rep.int(rgb(63, 63, 63, max = 255), 14))) +
  theme(legend.position = "none")


q
```


```{r}
ggsave(filename = "/Users/gemmaclucas/GitHub/Fecal_metabarcoding/Tern_pilot_study/Figures/18S_FOO_ROST_chicks_2017_2018_2panels.jpg",
  plot = q,
  dpi = 600,
  device = "jpeg",
  width = 5,
  height = 5
)
```

