---
title: "Pianka's Niche Overlap"
author: "Gemma Clucas"
date: "10/6/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(EcoSimR)
```

I'm following guidelines found in [this](https://cran.r-project.org/web/packages/EcoSimR/vignettes/nicheOverlapVignette.html) vignette. In there, they recommend the default algorithm ```ra3``` for shuffling row values to test significance.

### Read in the rarefied feature table

This is the same csv that I made for the ```FOO.Rmd``` script.

```{r data}
df <- read.csv("/Users/gemmaclucas/Dropbox/Diets_from_poop/2019_terns_puffins_fecal_data_analysis/MiFish/final_taxonomy_superblast/Terns/Terns_table_rarefied400_withtaxonomyedits.csv", header = TRUE) %>% 
  mutate(Year = as.factor(Year)) %>% 
  rename(River.herring = River.Herring,
         Sandlance = Sandlances,
         Mummichog = Mummichig)
```

## Niche overlap between Common Tern adults and chicks 
### Create matrix 

The function needs species (or in this case age groups) as rows and the prey items in columns. It will determine the proportion for each prey item, so it is fine to give it just the number of detections for each prey item for now.

I have lumped the data for 2017 and 2018 together, so that the test is just between adult and chick diets (years are not considered).

Select the data and calculate number of detections of each prey species:
```{r, message=FALSE, warning=FALSE }
COTE <- df %>% 
  dplyr::select(River.herring:White.hake, Species, Age) %>%    # keep just the columns of interest
  group_by(Species, Age) %>%                            
  mutate_if(is.numeric, ~1 * (. > 0)) %>%                           # change to detection/non-detection
  summarise_each(funs = sum) %>%                                    # sum up number of detections within in each group
  filter(Species == "COTE") %>%                                     
  mutate(group = paste(Species, Age, sep = "_")) %>%          # make new "group" name for each row
  ungroup() %>% 
  dplyr::select(group, River.herring:White.hake)

knitr::kable(COTE)  
```

### Calculate Pianka's niche overlap on the COTE matrix

To calculate just the score, you can use the ```pianka()``` function.
```{r}
COTE %>% dplyr::select(River.herring:White.hake) %>% 
  pianka()
```

To do simulations to caluclate significance, you can use the ```niche_null_model()``` function.
```{r}
COTE_pianka <- COTE %>% niche_null_model(nReps=5000, algo = "ra3", metric = "pianka")

summary(COTE_pianka)

```

Plot a histogram to see where the observed metric falls against the simulated data.
```{r}
plot(COTE_pianka, type = "hist")
```

This shows that there is much higher overlap than expected by chance. The observed overlap of `r COTE %>% dplyr::select(River.herring:White.hake) %>% pianka()` is well outside of the 95% confidence interval, which is denoted by the short dashed lines.

## Niche overlap between Common and Roseate tern chicks 

Create a table with the FOO for COTE and ROST chicks (years are grouped).

```{r, message=FALSE, warning=FALSE}
COTE_ROST <- df %>% 
  dplyr::select(River.herring:White.hake, Species, Age) %>%    # keep just the columns of interest
  group_by(Species, Age) %>%                            
  mutate_if(is.numeric, ~1 * (. > 0)) %>%                           # change to detection/non-detection
  summarise_each(funs = sum) %>%                                    # sum up number of detections within in each group
  filter(Species == "COTE" | Species == "ROST") %>% 
  filter(Age == "chick") %>% 
  mutate(group = paste(Species, Age, sep = "_")) %>%          # make new "group" name for each row
  ungroup() %>% 
  dplyr::select(group, River.herring:White.hake)

knitr::kable(COTE_ROST) 
```

Calculate the niche overlap and estimate 95% confidence interval, then plot.
```{r}
COTE_ROST_pianka <- COTE_ROST %>% niche_null_model(nReps=5000, algo = "ra3", metric = "pianka")

summary(COTE_ROST_pianka)

plot(COTE_ROST_pianka, type = "hist")
```

The observed niche overlap of `r COTE_ROST %>% dplyr::select(River.herring:White.hake) %>% pianka()` between Common and Roseate terns is outside the 95% confidence interval, and so despite differences in diets, they do not have a statistically differentiated niche according to Pianka's Niche Overlap.


### Check that it's working by simulating random data

```{r}
#create a random data set with uniform (0,1) values
myRandomData <- matrix(runif(300), nrow=30)

# run null model with pianka index and ra3, 5000 replications
myRandomModel <- niche_null_model(speciesData=myRandomData, 
                            algo="ra3", metric="pianka", 
                            suppressProg=TRUE,nReps=5000)

# print summary of model and plot histogram
summary(myRandomModel)
plot(myRandomModel,type="hist")
```

Yes, this time the simulated metric is inside the 95% CI.